---
title: "An introduction to R for phonetics students"
date: "`r format(Sys.time(), '%d-%b-%Y')`"
output:
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Instructions

Open the Rmarkdown file `Week8_Assignment.Rmd` in RStudio and read through it, filling in the code and answering the questions.

When you are done, change the code above from `eval = FALSE` to `eval = TRUE`, then click on the `Knit` button. This will generate an HTML file you can view with your browser.

If there are errors in your code, R will let you know. Make sure there are no errors before submitting.

Please, submit both the `.Rmd` file and the generated HTML file to TurnItIn.

# Getting data into R

Make sure you have attached the tidyverse packages before you continue.

```{r tidy}
library(tidyverse)
```

First, load the `hillenbrand.txt` file. These data are from the study of Hillenbrand et al. (1995), "Acoustic characteristics of American English vowels", *The Journal of the Acoustical Society of America* 97, 3099-3111, DOI: [10.1121/1.411872](https://doi.org/10.1121/1.411872).

Add the code to read the file and run it. You will need to specify this argument `delim = "\t"` in the function call.

```{r hill}
hillenbrand <- 
```

<!-- Hint: You may want to use `read_delim()`. -->

Let's inspect the data.

```{r hill-check}
hillenbrand
```

The first column, `MWBG`, indicates the speaker's age and gender. It is a factor with 4 levels (`b`, `g`, `m`, and `w`) which stand for "boy", "girl", "man", and "woman", respectively. 

Each row gives the average duration, f0, and first four formants for a given vowel quality for a given speaker.

# Numerical summaries

Each answer should have two parts: the R code you used to get the answer, and the answer itself. 

Give your answers with the appropriate scale (e.g., "ms" for duration). Round your answers to the nearest integer value.

1. What is the mean duration for the vowel `ae` over all speakers? (Get the mean duration for all vowels and report the mean duration of `ae`)

```{r}
hillenbrand %>%
  group_by(...) %>%
  summarise(
    duration_mean = ...
  )
```

2. What is the mean f0 for boys, averaging over all vowels?

```{r}
hillenbrand %>%
  group_by(...) %>%
  summarise(
    f0_mean = ...
  )
```

3. What is the mean f0 for the `aw` vowel for men?

```{r}
hillenbrand %>%
  group_by(..., ...) %>%
  summarise(
    ... = ...
  )
```

4. What is the median for F2 for the `uw` vowel for girls?

```{r}
hillenbrand %>%
  group_by(..., ...) %>%
  summarise(
    ... = ...
  )
```

5. What is the standard deviation for F1 for the `oo` vowel for women?

```{r}
hillenbrand %>%
  group_by(..., ...) %>%
  summarise(
    ... = ...
  )
```

6. Does the data set contain the same number of observations of the vowel `ei` for each age/gender combination?

```{r}
hillenbrand %>%
  group_by(....) %>%
  summarise(
    obs_n = ...
  )
```

7. Find the mean and standard deviation of the duration measures for each speaker age/gender group.

```{r}



```

8. Find the mean and standard deviation of f0 for each vowel.

```{r}



```


# Graphical summaries


9. Create a strip chart of F1 of each vowel, averaging over speaker age/gender. Make sure that the F1 values are plotted on the *y*-axis and that the *y*-axis is labelled `F1 (Hz)`.

```{r}
hillenbrand %>%
  ggplot(aes(..., ...)) +
  geom_...() +
  labs(
    ....
  )
```


10. Create a strip chart of vowel duration for each age/gender. Plot duration on the *x*-axis, and make sure the *x*-axis is labelled `Duration (ms)`. 

```{r}
hillenbrand %>%
  ggplot(aes(..., ...)) +
  ....
```

11. Reorder speaker age/gender so that the levels are in the order men - women - boys - girls. Then plot a strip chart of F1 for each vowel, faceted by speaker age/gender.

```{r}
hillenbrand %>%
  mutate(
    MWBG_re = factor(MWBG, levels = ....)
  ) %>%
  ggplot(aes(..., ...)) +
  .... +
  facet_grid(~ ...)
```

12. Plot histograms of f0, faceted by vowel and filled by speaker age/gender. Make sure the f0 variable is on the *x*-axis and that this is appropriately labelled. 

```{r}
hillenbrand %>%
  ggplot(aes(..., fill = ...)) +
  ... +                 # what's the right geometry?
  ...(~ ...) +          # what's the function for faceting?
  labs(
    ...
  )
```

13. Plot the f0 over all vowels for each speaker age/gender type as density plots. Instead of faceting, supply the call to the density geom with the argument `alpha = 0.5`. Make sure that the f0 variable is on the *x*-axis and that this is appropriately labelled.

```{r}
hillenbrand %>%
  ggplot(aes(..., fill = ...)) +
  ...(alpha = ...) +
  labs(
    ...
  )
```


14. What is the following plot suggesting about the relationship between F1 and duration (run the code to see the plot)? What do you think this means in articulatory terms?

```{r}
hillenbrand %>%
  ggplot(aes(F1, Duration)) +
  geom_point(aes(colour = Vowel)) +
  geom_smooth(method = "lm", formula = "y ~ x") +
  labs(
    x = "F1 (Hz)", y = "Duration (ms)"
  ) +
  theme_minimal()
```

<br>

<br>
